---
title: "Frequency EDA"
output: html_document
---

```{r}
# Load libraries
library(data.table)
library(here)
library(ggplot2)
library(ggrepel)

project.dir = here()

# Load data
dt = fread(here("data", "sim-modeling-dataset.csv"))

# Load helper functions
source(here("scripts", "EDA", "Frequency EDA Helper.R"))

# Global variables
cat_cols = c("year", "nb.rb", "driver.gender", "marital.status",
             "ncd.level", "region", "body.code", "seats", "fuel.type")

num_cols = c("driver.age", "yrs.licensed", "vehicle.age", "vehicle.value",
             "ccm", "hp", "weight", "length", "width", "height", "prior.claims")

cols = c(cat_cols, num_cols)
  
```

### Data Preparation
Turn categorical features into factors & Set base level to category with most exposures

```{r}
# Set categorical data as factors
dt = set_categorical_features(dt)

# Set levels for categorical data
dt = set_categorical_levels(dt, cat_cols)

```

### One-Way Frequency Models
I plotted the AIC & Deviance of every possible one-way model, including the null model which is just the intercept. 
Each one-way model is a Poisson regression model with claim count as the dependent variable and log(exposure) as the offset
Features to the bottom left of the figure have more predictive power, so I'll prioritize these features when modeling. 
These features include: ncd.level, year, yrs.licensed, nb.rb, region, prior.claims, vehicle.age, driver.age, height, body.code, marital.status

```{r fig.width=15, fig.height=8}

# one way frequency models
one_way_dt = compute_frequency_aic_dev(dt, vars=cols)

# Plot the AIC-deviance plot
plt = ggplot(one_way_dt, aes(x = AIC, y = Deviance, label = vars)) +
      geom_point() +
      geom_text_repel(max.overlaps = Inf)
show(plt)

```
### Numeric Features EDA

yrs.licensed: 

There's a clear negative relation between frequency and yrs.licensed. 
Yrs.licensed values (2, 4] have negligible differences in frequency so they can be binned together.
Yrs.licensed values (5,10] have negligible differences in frequency so they can be binned together. 

New features for frequency modeling will be labeled with a ".f" 
yrs.lic.f: new features with breaks ->
  (1], (2], [3, 4], 5, 6+


```{r}
# numeric frequency relationship 
col = "yrs.licensed"
freq = numeric_freq_relation(col=col, dt = dt, n_bins = 10, bin_data = FALSE)

# construct new feature for modeling
dt = feature_engineer_numeric(dt, col="yrs.licensed", new_col="yrs.licensed.f", breaks=c(0, 1, 2, 4, 5, Inf), labels=c("1", "2", "[3, 4]", "5", "6+"))







```

```{r}

# Bin data s.t. exposures are approximately even across bins
x_col = "yrs.licensed"
n_bins = 5
dt_exp = dt[, .(exp=sum(exposure)), by = .(x = get(x_col))][order(x)]
total_exp = sum(dt_exp$exp)
target_exp = total_exp/n_bins
dt_exp[, cumexp := cumsum(exp)]
dt_exp[, bin_id := pmin(n_bins, ceiling(cumexp/target_exp))]

# Create bin edges
bin_min = dt_exp[, .(left_bin=min(x)), .(bin_id)]
bin_max = dt_exp[, .(right_bin=max(x)), .(bin_id)]
bin_edges = merge(bin_min, bin_max, by="bin_id")
breaks = c(min(bin_edges$left_bin) - 1, sort(unique(bin_edges$right_bin)))


```





```{r fig.width=15, fig.height=8}
# numeric frequency relationship 
col = "driver.age"
freq = numeric_freq_relation(col=col, dt = dt, n_bins = 30, bin_data = FALSE)
freq = numeric_freq_relation(col=col, dt = dt, n_bins = 15, bin_data = TRUE)

```

dta$drv.age.gr1[dta$driver.age %in% 18:33] <- "18-33"
dta$drv.age.gr1[dta$driver.age %in% 34:38] <- "34-38"
dta$drv.age.gr1[dta$driver.age %in% 39:43] <- "39-43"
dta$drv.age.gr1[dta$driver.age %in% 44:50] <- "44-50"
dta$drv.age.gr1[dta$driver.age %in% 51:60] <- "51-60"
dta$drv.age.gr1[dta$driver.age %in% 61:69] <- "61-69"
dta$drv.age.gr1[dta$driver.age %in% 70:93] <- "70+"




